abxD01 Paper

### 2015-03-09 PAPER.Rmd
Update discussion

-community effect—have other things to compensate—highlight examples –streptomycin blue triangles—where low lachno but no cdiff so what else is protecting? Not one otu that explains it all… just cus low levels doesn’t screw you, also see recovered points in 5C-angela, sonnenburg-need other thingsssss
  -fig2—there are like larger groups in protection/susceptibility
	-combo of loss of good bugs, and gain of bad—a model of infection
-tie back to mbio paper, with enterobacteriaceae they are screwed
-modeling—not much done for CDI, link back to humans, shows we could model risk as they come in—put towards end with the larger picture
We designed statistical models based on a data set of varied antibiotic perturbations of the murine microbiota and their subsequent CR levels against *C. difficile*. Though antibiotic administration induced unique microbiota structures, colonization resistance was similarly low against *C. difficile*.
These models incorporated all combinations of top candidate species-level OTUs. We then tested our top model’s ability to use new abundance information on a specific subset of the microbiota to predict the subsequent *C. difficile* colonization level, which was experimentally validated. These results greatly expand our knowledge of various perturbations and their subsequent effects on microbiota structure and function (colonization resistance ability). Furthermore modeling important/key bacterial subsets of the community we can build models to assess CDI risk.

@@

Furthermore, the dynamics of *C. difficile* colonization does not appear to be dependent soley on the loss of protective species. Both *Escherichia* OTU11 and *Streptococcus* OTU78 were positively correlated with *C. difficile* colonization and important in the random forest model. Notably, *Escherichia* OTU11 was among the strongest positively correlated OTUs with *C. difficile* across all antibiotic treatments, except for streptomycin. In fact, while streptomycin treatment (5 mg/ml) lead to high *C. difficile* colonization, these communities had very little, if any, *Escherichia* OTU11. This demostrates that while *Escherichia* may aid in colonization of *C. difficile*, it is not essential. These observations lead to a model of colonization by *C. difficile* in which antibiotic perturbation leads to loss of protective species.

@@

In the context of antibiotic perturbation however, widespread loss of protective bacteria allows the growth of bacteria which otherwise are outcompeted, including many *Proteobacteria* (*E. coli*).

Furthermore, the ability of *C. difficile* to colonize a community does not appear to be dependent soley on the loss of protective species.  Rather, it seems likely that as protective species are lost, advantageous bacteria such as *E. coli* can colonize quickly and adapt, which in turn also promotes a favorable environment for *C. dfficile* colonization. Many members of the proteobacteria---attributes: grow quickly, wide array of nutrient sources, more generalists. Examples of e.coli causing inflammation, which can help pathogens in some instances where bad for resident good bugs. Also e. coli using nutrients within mucus (?) or at least contributing to loss of oxygen near epithelium, which can also let pathogens get closer and colonize.. cause further damage (by toxins, etc)...

>>>>

-community effect—have other things to compensate—highlight examples –streptomycin blue triangles—where low lachno but no cdiff so what else is protecting? Not one otu that explains it all… just cus low levels doesn’t screw you, also see recovered points in 5C-angela, sonnenburg-need other thingsssss
-tie back to mbio paper, with enterobacteriaceae they are screwed
-modeling—not much done for CDI, link back to humans, shows we could model risk as they come in—put towards end with the larger picture
We designed statistical models based on a data set of varied antibiotic perturbations of the murine microbiota and their subsequent CR levels against *C. difficile*. Though antibiotic administration induced unique microbiota structures, colonization resistance was similarly low against *C. difficile*.
These models incorporated all combinations of top candidate species-level OTUs. We then tested our top model’s ability to use new abundance information on a specific subset of the microbiota to predict the subsequent *C. difficile* colonization level, which was experimentally validated. These results greatly expand our knowledge of various perturbations and their subsequent effects on microbiota structure and function (colonization resistance ability). Furthermore modeling important/key bacterial subsets of the community we can build models to assess CDI risk.

@@

Furthermore, the dynamics of *C. difficile* colonization does not appear to be dependent soley on the loss of protective species. Both *Escherichia* OTU11 and *Streptococcus* OTU78 were positively correlated with *C. difficile* colonization and important in the random forest model. Notably, *Escherichia* OTU11 was among the strongest positively correlated OTUs with *C. difficile* across all antibiotic treatments, except for streptomycin. In fact, while streptomycin treatment (5 mg/ml) lead to high *C. difficile* colonization, these communities had very little, if any, *Escherichia* OTU11. This demostrates that while *Escherichia* may aid in colonization of *C. difficile*, it is not essential. These observations lead to a model of colonization by *C. difficile* in which antibiotic perturbation leads to loss of protective species. This in turn eliminates competition, frees resources, and eliminates direct or indirect inhibitory effects of bacterial products (e.g. bacteriocins or short chain fatty acids) made by these protective bacteria. While this may be suitable enough for *C. difficile* to colonize, other opportunistic species such as *Escherichia coli* and other *Proteobacteria* may also take advantage of these microbiota perturbations. Increased colonization by these bacteria can increase inflammation, produce a new profile of metabolites and bacterial products, reduce the mucus layer,  and deoxygenate the microoxic tiers nearing the epithelium. These changes in the gut milieu may be advantageous for *C. difficile* and other gut pathogens. Inflammation, for instance, has been shown to be advantageous for *Salmonella tyhpumurium* to further disrupt the indigenous microbiota []. Inflammation also leads to a shift in the availability of resources, such as nitrites, which can be used as an alternative resource by *Salmonella typhimurium* []. *E.coli* is capable of using resources found in the mucus layer while simultaneously deoxygenating this region []. This reduces the spatial protection of the mucus layer, allowing bacteria, including strictly anaerobic bacteria, and bacterial products (e.g. toxins) to more readily reach the epithelial layer.

@@
